---
title: "P8105 Homework 2"
author: "Ruiyang Li"
output: github_document
---

This is my solution to HW2.

```{r setup}
library(readxl)
library(tidyverse)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet:

```{r}
mr_trash_wheel = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "Mr. Trash Wheel", 
             range = cell_cols("A:N"),
             skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls, 0))
  ) 
```


Read and clean precipitation data for 2017 and 2018:

```{r}
precip_2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "2017 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names() %>%   
  mutate(year = "2017") %>% 
  relocate(year)
  
precip_2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
             sheet = "2018 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2018") %>% 
  relocate(year)
```

Combine precipitation datasets:

```{r}
precip = 
  bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month])
```


About these data: 

For the Mr. Trash Wheel dataset, there are `r ncol(mr_trash_wheel)` variables with `r nrow(mr_trash_wheel)` observations. Variables include: `r colnames(mr_trash_wheel)`. Data on `r length(unique(pull(mr_trash_wheel, dumpster)))` dumpsters were collected from `r min(pull(mr_trash_wheel, year))` to `r max(pull(mr_trash_wheel, year))`. The total weight of trash is `r sum(pull(mr_trash_wheel, weight_tons))` tons. These trash was incinerated to provide electricity for `r format(round(sum(pull(mr_trash_wheel, homes_powered)), 0), scientific = F)` Maryland homes. The median number of sports balls in a dumpster is `r median(pull(mr_trash_wheel, sports_balls))`.

For the combined precipitation dataset, there are `r ncol(precip)` variables with `r nrow(precip)` observations. Variables include: `r colnames(precip)`. For example, the total precipitation in 2018 was `r sum(pull(precip_2018, total))` (in). 



## Problem 2 

Read and clean the nyc transit data: 

```{r}
nyc_transit = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

About this dataset: 

The raw NYC transic dataset contains `r ncol(read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"))` variables. During the data cleaning process, I first cleaned the variables' names by making them all snake case, then I selected the `r ncol(nyc_transit)` variables of interest: `r colnames(nyc_transit)`, and finally I converted the variable `entry` from character (`YES` vs `NO`) to logical (`TRUE` vs `FALSE`). The resulting dataset has `r ncol(nyc_transit)` variables with dimension `r nrow(nyc_transit)` x `r ncol(nyc_transit)`. Data in this resulting dataset are not tidy. For example, the current 11 route variables can be reduced into one route variable indicating different route numbers. 




